---
Decision Tree and Cross Validation 
title: "Assignment 6"
author: "Juan Djuwadi"
date: "11/16/2016"
output: html_document
---
#Addignment 6

In this assignment you will be looking at data from a MOOC. It contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  
forum.posts (numeric) - How many forum posts a student made throughout the course  
grade (numeric) - A student's average grade for the course exam  
assignment (numeric) - A student's average grade for the course assignments  

#Packages
```{r}
library(rpart)
```

#Data
```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv", header = TRUE)

M2 <- read.csv("MOOC2.csv", header = TRUE)

```

#Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame. Which variables should you use?

#Count the Number of Yes in the M1 Column
sum(as.numeric(M1$certified == "yes"))
#725 is the amount that is certified 
#275 us the amount that is not-certified

c.tree1 <- rpart(certified ~ grade + assignment, method = "class", data = M1)
c.tree2 <- rpart(certified ~ forum.posts + grade, method = "class", data = M1)
c.tree3 <- rpart(certified ~ forum.posts + assignment, method = "class", data = M1)

c.tree4 <- rpart(certified ~ scale(forum.posts) + grade + assignment, method = "class", data = M1, control = rpart.control(minsplit = 2))
#We set minsplit = 2 to indicate the minimum numer of nodes that the tree must branch out of. Forum posts are scaled.Here we will be using this variable and settings to evaluate the rest of the data. 

c.tree41 <- rpart(certified ~ scale(forum.posts) + assignment, method = "class", data = M1, control = rpart.control(minsplit = 2))

c.tree5 <- rpart(certified ~ forum.posts, method = "class", data = M1)
c.tree6 <- rpart(certified ~ assignment, metho = "class", data = M1)

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)
printcp(c.tree2)
printcp(c.tree3)
printcp(c.tree4)
printcp(c.tree41)
printcp(c.tree5)
printcp(c.tree6)
printcp(c.treeTree)

#Plot your tree

post(c.tree1, file = "tree1.ps", title = "MOOC") #This creates a pdf image of the tree
post(c.tree2, file = "tree2.ps", title = "MOOC")
post(c.tree3, file = "tree3.ps", title = "MOOC")
post(c.tree4, file = "tree4.ps", title = "MOOC")
post(c.tree41, file = "tree41.ps", title = "MOOC")
post(c.tree5, file = "tree5.ps", title = "MOOC")
post(c.tree6, file = "tree6.ps", title = "MOOC")
post(c.treeTree, file = "treeTree.ps", title = "MOOC") 
```

#The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation. CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions. However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

#If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

```{r}
c.treePrune <- prune(c.tree4, cp = 0.5)#Set cp to the level at which you want the tree to end

#We set the CP = 0.5 to indicate that we only want one node in the graph thus pruning the second node. 

#Visualize this tree and compare it to the one you generated earlier

post(c.treePrune, file = "treePruned.ps", title = "MOOC") #This creates a pdf image of the tree

#Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

M2$predict1 <- predict(c.tree4, M2, type = "class")

M2$predict2 <- predict(c.treePrune, M2, type = "class")

printcp(c.tree4)
printcp(c.treePrune)

#We can see from the error results that both the CP and error rate results are the same. By extension, both trees are similar. 

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)

```
